The World Bank, Women, Business and the Law

### 2024-02-20 outputs/paper/paper.qmd
Adding workplace and pay

# Create a data frame for the table
data <- data.frame(
  Indicator = "Pay",
  Score = 100,
  Question = "“Does the law mandate equal remuneration for work of equal value?”",
  Legal_Basis = "Ontario Pay Equity Act, Sec. 6",
  Achievement_of_pay_equity = "6 (1) For the purposes of this Act, pay equity is achieved under the job-to-job method of comparison when the job rate for the female job class that is the subject of the comparison is at least equal to the job rate for a male job class in the same establishment where the work performed in the two job classes is of equal or comparable value. R.S.O. 1990, c. P.7, s. 6 (1); 1993, c. 4, s. 4 (1).",
  Relevant_article = "The gender wage gap in Canada: 1998 to 2018 C Labour Statistics: Research Papers by Rachelle Pelletier, Martha Patterson and Melissa Moyser",
  Stylized_fact = "Women still earn 87 cents for every dollar a man earns in Canada",
  Link_to_Article = "[Read more](https://www150.statcan.gc.ca/n1/en/pub/75-004-m/75-004-m2019004-eng.pdf?st=szPFyv6s)",
  Database = "Annualized data from the Labour Force Survey (LFS), a monthly household survey conducted by Statistics Canada.",
  Additional_info = "56,000 households, no aboriginal settlements or full-time Canadian Armed Force members, remote households in low population density areas"
)

# Transpose the data frame
data <- t(data)

# Create a table using kableExtra
kable(data, format = "markdown") %>%
  kable_styling()
```

## Second discussion point {#sec-point2}

## Third discussion point {#sec-point3}

>>>>

# Create a data frame for the table
pay <- data.frame(
  Indicator = "Pay",
  Score = 100,
  Question = "“Does the law mandate equal remuneration for work of equal value?”",
  Legal_Basis = "Ontario Pay Equity Act, Sec. 6",
  Achievement_of_pay_equity = "6 (1) For the purposes of this Act, pay equity is achieved under the job-to-job method of comparison when the job rate for the female job class that is the subject of the comparison is at least equal to the job rate for a male job class in the same establishment where the work performed in the two job classes is of equal or comparable value. R.S.O. 1990, c. P.7, s. 6 (1); 1993, c. 4, s. 4 (1).",
  Relevant_article = "The gender wage gap in Canada: 1998 to 2018 C Labour Statistics: Research Papers by Rachelle Pelletier, Martha Patterson and Melissa Moyser",
  Stylized_fact = "Women still earn 87 cents for every dollar a man earns in Canada",
  Link_to_Article = "[Read more](https://www150.statcan.gc.ca/n1/en/pub/75-004-m/75-004-m2019004-eng.pdf?st=szPFyv6s)",
  Database = "Annualized data from the Labour Force Survey (LFS), a monthly household survey conducted by Statistics Canada.",
  Additional_info = "56,000 households, no aboriginal settlements or full-time Canadian Armed Force members, remote households in low population density areas"
)

# Transpose the data frame
pay <- t(pay)

# Replace underscores with spaces in column names
colnames(pay) <- gsub("_", " ", colnames(pay))

# Create a table using kableExtra
  kable(pay, format = "latex", caption = "Summary of findings -Pay Indicator") %>%
  column_spec(2, width = "27em")

The LFS survey reveals that in 2018, female employees aged 25 to 34 earned $4.13 less per hour compared to their male counterparts, equaling 87 cents on the dollar. In the decade between 1988 and 2018, the gap has reduced by merely $1.04. 
Given that women in Canada are now better educated on average than men, acquiring bachelor’s degrees at a faster pace, and have diversified their fields of study, particularly in STEM, the persistence of this gap warrants a closer look.

The authors used a Blinder Oaxaca Decomposition to quantify how much of the wage difference can be explained by control factors. Some of these factors included level of education, job tenure, union coverage, and demographics. The results show that even after controlling for the above factors, roughly two-thirds of the wage gap is unexplained, and the difference is statistically significant at a 95% confidence interval.

The table below reproduces the gender wage gap- Oaxaca decomposition data from the Labour Statistics report. As is visible, only an additional 7% of this gap has become explicable in the years that have followed since 1988, largely due to unionization reducing men's wages, and women entering new industries. Despite extensive research, the gender wage gap remains an unsolved mystery in Canada.

```{r, results='asis', echo=FALSE, fig.cap="Explanation of the Gender Wage Gap, 1998 to 2018"}

library(kableExtra)

# Create a data frame for the table
wagegap <- data.frame(
  Category = c("Total explained portion (sum of effects of variables below)", "Human capital", "Education", 
               "Job tenure", "Job attributes", "Part-time", "Public sector", "Union", "Firm size", 
               "Occupation and industry", "Occupation", "Industry", "Demographics", "Age", 
               "Province", "Children", "Marital status", "Total unexplained portion"),
  'In 1998' = c(28.0, 1.8, -0.5, 2.3, 10.1, 8.9, -0.6, 0.7, 1.1, 18.3, 1.8, 16.5, -2.2, -0.4, -0.1, -0.3, -1.4, 72.0),
  'In 2018' = c(36.6, -6.1, -4.8, -1.3, -0.7, 9.2, -5.3, -3.4, -1.1, 44.8, 5.1, 39.7, -1.6, -0.3, 1.0, -0.8, -1.5, 63.4)
)

# Replace special characters in column names
#colnames(wagegap) <- gsub("`", "", colnames(wagegap))

# Create a table using kableExtra
kable(wagegap, format = "latex", booktabs = TRUE, escape = FALSE) %>%
 kable_styling() %>%
#  add_header_above(c("Category" = 1, "1998" = 1, "2018" = 1)) %>%
  row_spec(18, bold = TRUE)

## Workplace {#sec-point2}

The next indicator under scrutiny is the Workplace, with a specific focus on legislation addressing sexual harassment. While legal support is evident in various codes within this category, it is essential to emphasize that the mere existence of legal provisions does not guarantee their effective implementation. The following analysis then, acknowledges the potential disparity between legal frameworks and their real-world impact on preventing and addressing sexual harassment in the workplace.

# Create a data frame for the table
workplace <- data.frame(
  Indicator = "Workplace",
  Score = 100,
  Question = "“Is there legislation on sexual harassment in employment??”",
  Legal_Basis = "Canadian Human Rights Act, Sec. 14(1) and (2); Ontario Occupational Health and Safety Act, Sec. 1 and 32.0.1; Ontario Human Rights Code, Sec. 7(2) and 7(3)",
  Code = "14 (1) It is a discriminatory practice, in matters related to employment,
to harass an individual on a prohibited ground of discrimination.Without limiting the generality of subsection 1, sexual harassment shall, for the purposes of that subsection, be deemed to be harassment on a prohibited ground of discrimination.",
  Relevant_article = "Gender Results Framework: A new data table on workplace harassment",
  Stylized_fact = "Women still earn 87 cents for every dollar a man earns in Canada",
  Link_to_Article = "[Read more]https://www150.statcan.gc.ca/n1/daily-quotidien/240212/dq240212a-eng.htm",
  Database = "Data from 2020 Survey on Sexual Misconduct at Work (SSMW) and Gender Results Framework",
  Additional_info = "56,000 households, no aboriginal settlements or full-time Canadian Armed Force members, remote households in low population density areas"
)

# Transpose the data frame
workplace <- t(workplace)

# Replace underscores with spaces in column names
colnames(workplace) <- gsub("_", " ", colnames(workplace))

# Create a table using kableExtra
  kable(workplace, format = "latex", caption = "Summary of findings -Pay Indicator") %>%
  column_spec(2, width = "27em")

Utilizing data from Statistics Canada's Centre for Gender, Diversity, and Inclusion Statistics, in collaboration with the Canadian Centre for Justice and Community Safety Statistics, we analyze a recently released data table on workplace harassment and sexual assault experiences. The table reports on the Gender Results Framework (GRF), initiated by the Government of Canada in 2018 to monitor gender equality progress. Additionally, the data incorporates insights gathered from the 2020 Survey on Sexual Misconduct at Work (SSMW).

The reproduced table from the 2020 Survey on Sexual Misconduct at Work outlines the proportions of employees reporting inappropriate sexualized behavior, discriminatory behavior, and sexual assault in their work environments. Notably, the data highlights the heightened vulnerability of individuals in the 25 to 34 age group, where 57% of females report experiencing sexualized behaviors. Within this age group, women are twice as likely as men to face discrimination and three times as likely to experience sexual assault.
```{r}
# Install and load necessary packages if not already installed
 # install.packages("tidyverse")

# Install and load necessary packages if not already installed
if (!requireNamespace("tidyverse", quietly = TRUE)) {
  install.packages("tidyverse")
}

# Load the necessary libraries
library(tidyverse)

# Your provided data
data <- read.table(text = '"Group" "M - inappropriate sexualized behaviours" "M - Discrimination" "M - SA" "F - inappropriate sexualized behaviours" "F - Discrimination" "F - SA"
"Total, all people" 29 9 3 44 20 13
"Age 15 to 24" 27 12 0 48 25 7
"Age 25 to 34" 37 13 5 57 27 16
"Age 35 to 44" 34 11 3 41 20 12
"Age 45 to 54" 28 8 3 42 18 12
"Age 55 and older" 24 6 2 39 16 15', header = TRUE)

# Transform the data into long format for ggplot
data_long <- data %>%
  pivot_longer(cols = -Group, names_to = "Category", values_to = "Percentage")

# Create a sideways stacked bar plot
ggplot(data_long, aes(x = Percentage, y = Group, fill = Category)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.7) +
  scale_fill_manual(values = c("#FF9999", "#FF6666", "#FF3333", "#66B2FF", "#3366FF", "#000099"), name = "Category") +
  labs(title = "Sideways Stacked Bar Plot",
       x = "Percentage",
       y = "Group") +
  theme_minimal() +
  theme(
    axis.text.y = element_text(angle = 0, hjust = 1),
    legend.position = "bottom",
    legend.direction = "horizontal",
    legend.box = "horizontal",
    legend.margin = margin(t = 0, unit = "pt")
)

Further analysis of Canadian data reveals that nearly half of women, at 47%, report experiencing some form of harassment or sexual assault in the workplace. The Gender Results Framework (GRF), a key tool for monitoring gender equality in Canada, indicates that women are twice as likely as men to encounter unwanted sexual behaviors in a work environment. Beyond the workplace, Statistics Canada notes that women are three times more likely to experience unwanted sexual behaviors in public, four times more likely to have been sexually assaulted since age 15, and twice as likely to suffer negative emotional consequences from such experiences compared to men. 
These findings underscore the pervasive nature of gender-based harassment and assault, both within and outside professional settings in Canada, which are far from a pristine perfect ‘100'.
